Modeling Macroeconomic Phenomena \\with Multi-Agent Reinforcement Learning

### 2019-12-06 paper/paper.tex
Andy's edits to paper

While there are relatively few examples of reinforcement learning applied to macroeconomic models, there are quite a few papers with relevant material on applying multi-agent models to various settings involving competition and collaboration. For example, a 2013 study by Bos et. al. found that reinforcement learning alone does not accurately model humans in competitive social environments, but it \textit{does} accurately model this when combined with biased reward representations \cite{vandenBos2137}. In a separate work, Bereby-Meyer and Roth experiment with reinforcement learning in game theoretic situations, such as the repeated prisoner's dilemma, when complicated by noise \cite{10.1257/aer.96.4.1029}. The authors conclude that noise slows learning and strongly negatively affects cooperative tendencies \cite{10.1257/aer.96.4.1029}. As such, even with simple multi-agent situations involving some randomness, strong cooperation is unlikely.

@@

The space of all actions is vast. In order to make the problem tractable, the offered price and offered units are discretized into $P$ and $U$ values, respectively. In our simulations, we often set $P \approx 15$ and $U \approx 25$. Similarly, the demand curve was parameterized by families of functions. We experimented with two different families of curves: inverse and linear. The family of inverse curves $y = a/x + b$ is specified by the two parameters $a, b$ (in our simulations we used approximately 36 total possible configurations). The family of linear curves was specified by solely the $y$-intercept, as the $x$-intercept was a hyperparameter fixed before training (the $x$-intercept corresponds to the maximum number of goods that can be purchased or produced by a given human or firm in a single iteration). In practice, the linear family trained faster, but the inverse family led to better results. We use the family of inverse demand curves for all plots presented in this paper.

@@

We implemented the circular flow model using multi-agent reinforcement learning and demonstrated that the results were qualitatively correct and could be used to investigate more complex macroeconomic phenomena. In particular, we found reasonable results in a simple monopolization model with a single firm agent and five human agents. This model was tested with two different policy gradient algorithms: REINFORCE and Q-actor critic. The latter was faster, but suffered from stability issues as compared to the former. Additionally, we simulated a more complex situation involving 5 firm agents and 25 human agents. The resulting dynamics had realistic features, including policies that corresponded to retirement behavior, slowly increasing wealth inequality, etc.

>>>>

While there are relatively few examples of reinforcement learning applied to macroeconomic models, there are quite a few papers with relevant material on applying multi-agent models to various settings involving competition and collaboration. For example, a 2013 study by Bos et. al. found that reinforcement learning alone does not accurately model humans (who deviate from rational behavior) in competitive social environments, but it \textit{does} accurately model this when combined with biased reward representations \cite{vandenBos2137}. In a separate work, Bereby-Meyer and Roth experiment with reinforcement learning in game theoretic situations, such as the repeated prisoner's dilemma, when complicated by noise \cite{10.1257/aer.96.4.1029}. The authors conclude that noise slows learning and strongly negatively affects cooperative tendencies \cite{10.1257/aer.96.4.1029}. As such, even with simple multi-agent situations involving some randomness, strong cooperation is unlikely.

@@

The space of all actions is vast. In order to make the problem tractable, the offered price and offered units are discretized into $P$ and $U$ values, respectively. In our simulations, we often set $P \approx 15$ and $U \approx 25$. Similarly, the demand curve was parameterized by families of functions. We experimented with two different families of curves: inverse and linear. The family of inverse curves $y = a/x + b$ is specified by the two parameters $a, b$ (in our simulations we used approximately 36 total possible configurations). The family of linear curves was specified by solely the $y$-intercept, as the $x$-intercept was a hyperparameter fixed before training (the $x$-intercept corresponds to the maximum number of goods that can be purchased or produced by a given human or firm in a single iteration). In practice, the linear family trained faster, but the inverse family led to more realistic behavior. We use the family of inverse demand curves for all plots presented in this paper.

@@

We implemented the circular flow model using multi-agent reinforcement learning and demonstrated that the results were qualitatively correct and could be used to investigate more complex macroeconomic phenomena. In particular, we found reasonable results in a simple monopolization model with a single firm agent and five human agents. This model was tested with two different policy gradient algorithms: REINFORCE and actor critic. The latter's policy gradients converged faster, but the overall algorithm suffered from stability issues as compared to the former. Additionally, we simulated a more complex situation involving 5 firm agents and 25 human agents. The resulting dynamics had realistic features, including policies that corresponded to retirement behavior, slowly increasing wealth inequality, etc.
